---
title: "ESTP training course"
output:
  html_notebook: default
  pdf_document: default
  word_document: default
---

The required packages.
```{r}
library(RJDemetra)
library(JDLight)
```

Load two time-series from JDLight.

```{r}
a<-retail$HobbyToyAndGameStores
b<-retail$JewelryStores
```

# Preliminary analysis of the series

```{r}
agg<-a+b
plot(ts.union(a, b, agg))
```
We can notice the same pattern. E.g. the spikes in December. In addition we observe a spike in December 1996 in series a. However we do not notice the spike in the aggregate. 

# Indirect Seasonal Adjusment
```{r}
a_x13<-x13_def(a,"RSA5")
a_sa<-a_x13$final$series[,"sa"]
b_x13<-x13_def(b,"RSA5")
b_sa<-b_x13$final$series[,"sa"] 
```
We use X13 to seasonally adjust the two series. 
```{r}
agg_sa_ind<-a_sa+b_sa
plot(ts.union(a_sa, b_sa, agg_sa_ind))
```
We built the seasonally adjusted aggreagate. 
In the seasonal adjusted figure we can clearly see the outlier in December 1996. In the aggregate it is not noticeable in the same extent.
```{r}
plot(agg_sa_ind)

```

We want to test for seasonality. We see that the series neither displays seasonality for the QS Method nor the Friedman method. 
```{r}
jd_seasonality(agg_sa_ind, test="QS")
jd_seasonality(agg_sa_ind, test="FRIEDMAN")
```

# Direct Seasonal Adjustment
```{r}
agg_sa_dir_x13<-x13_def(agg,"RSA5")
agg_sa_dir<-agg_sa_dir_x13$final$series[,"sa"]
ts.plot( agg_sa_dir)
```
Again we check whether seasonaltiy remains in the seasonal adjusted series. 
We see there is no seasonality left. 
```{r}
jd_seasonality(agg_sa_dir, test="QS")
jd_seasonality(agg_sa_dir, test="FRIEDMAN")

```
# Comparing the results of the two methods
```{r}
ts.plot(ts.union(agg_sa_dir, agg_sa_ind), col=c("black", "red"))
```
The indirect method captures the downhill peak while the direct method eliminated it. We assume this is due to the outlier detection. 




#### Outliers of series a
```{r}
a_x13$regarima$regression.coefficients

```

#### Outliers of series b
```{r}
b_x13$regarima$regression.coefficients

```
#### Outliers of the aggregated series
```{r}
agg_sa_dir_x13$regarima$regression.coefficients
```
We notice that the two outliers May and December 1996 are found in series a but not in the aggregate of series a and b. However we find a new outlier in September 2008. 

We can conclude that in this particular case direct seasonal adjustment make the outlier in May 1996 disappear.

```{r}
data.frame(month=1:12, 
           direct=window(agg_sa_dir, c(1996,1), c(1996,12)), 
           indirect=window(agg_sa_ind, c(1996,1), c(1996,12)))
```
## Anaylsis of the differences
```{r}
dif<-agg_sa_dir-agg_sa_ind
plot(dif)

```
We can see seasonality in the difference of the two different seasonal adjusted aggregates.
```{r}
jd_seasonality(dif)
```
This is confirmed by the test.
So, why is that?
